### 2021-12-29 cleave.tex
writing

Once the core components for the emulation have been defined and implemented, the emulation needs to be configured to register them with the framework.
This is done through fully-featured Python scripts which simply define a number of required (and some optional) top-level variables.
These Python configuration files are then passed as arguments to the \texttt{cleave} bash command provided by the framework, which sets up and executes the emulation.
The use of Python scripts for configuration allows for a high degree of flexibility and complexity in the definition of emulation setups.

At this point, yet another core component of CLEAVE needs to be mentioned: the Dispatcher server.
Although Controllers in CLEAVE can be started manually, this quickly becomes cumbersome when performing experimentation with more than a few Plants.
The CLEAVE Dispatcher server solves this by providing a way for remote deployment of Controllers through a RESTful HTTP service that the Plant connects to (see \cref{fig:cleave:dispatch}).

Example configuration files for a Plant and the Dispatcher are presented in \cref{lst:config:plant,lst:config:dispatcher}.

\begin{figure}
    \centering
    \includegraphics[width=.85\columnwidth]{images/CLEAVE_dispatcher.png}
    \caption{
        Diagram of the interaction between CLEAVE Plant, Dispatcher and Controller.
        The Dispatcher is instantiated and listens on a specific port on the Controller Host. 
        As the Plant initializes its emulation, it requests a Controller from the Dispatcher; the Dispatcher spawns the corresponding Controller and returns the host and port on which it is listening.
        The Plant proceeds to execute its emulation configuration.
        Finally, on shutdown, the Plant once again communicates with the Dispatcher to request the shutdown of the associated Controller.
    }\label{fig:cleave:dispatch}
\end{figure}

@@

host = '0.0.0.0'  # address of dispatch server
port = 8080       # providing Controllers

# Controller parameters
controller_class = 'InvPendulumController'
controller_params = {'ref': 0.0}

@@

actuators = [
    SimpleConstantActuator(
        initial_value=0,
        prop_name='force')
]
    \end{minted}
    \caption{
        Example configuration file for a CLEAVE Plant, defining the parameters for a single emulation.
        The ``controller parameters'' define the class of Controller to request from the Dispatcher.
        ``Plant emulation parameters'' define key emulation parameters, such as the emulation discrete-time update rate (the \emph{tick rate}).
        The Sensors and Actuators are then provided in lists, each Sensor and Actuator attached to a specific property identified by a string.
        Additionally, Sensors have a sampling rate in Hz which needs to be provided as an integer.
    }
    \label{lst:config:plant}
\end{listing}

\begin{listing}
    \begin{minted}[linenos=true,
        breaklines, 
        breakafter=d,
        fontsize=\small]{Python}
from cleave.impl.inverted_pendulum import InvPendulumController

host = 'localhost'
port = 8080

controllers = {
    'InvPendulumController': InvPendulumController,
}

output_dir = './controllers'
    \end{minted}
    \caption{
        Example configuration file for the CLEAVE Dispatcher.
        The previously defined Controller is registered with the framework and will then be available from the Dispatcher server for use with a Plant.
    }
    \label{lst:config:dispatcher}
\end{listing}

>>>>

Once the core components for the emulation have been defined and implemented, the emulation needs to be configured to register them with the framework.
This is done through fully-featured Python scripts which simply define a number of required (and some optional) top-level variables.
An example of such a configuration file can be seen in \cref{lst:config:plant}.
These Python configuration files are then passed as arguments to the \texttt{cleave} bash command provided by the framework, which sets up and executes the emulation.
The use of Python scripts for configuration allows for a high degree of flexibility and complexity in the definition of emulation setups.

@@

@@

actuators = [
    SimpleConstantActuator(
        initial_value=0,
        prop_name='force')
]
    \end{minted}
    \caption{
        Example configuration file for a CLEAVE Plant, defining the parameters for a single emulation.
        ``Plant emulation parameters'' define key emulation parameters, such as the emulation discrete-time update rate (the \emph{tick rate}).
        The Sensors and Actuators are then provided in lists, each Sensor and Actuator attached to a specific property identified by a string.
        Additionally, Sensors have a sampling rate in Hz which needs to be provided as an integer.
    }
    \label{lst:config:plant}
\end{listing}

### 2021-12-29 experiments.tex
writing

>>>>

\todo[inline]{WIP}

In this section, we aim to validate the utility of the CLEAVE framework through a series of experiments.

\todo[inline]{talk about goal of experiments}

Our experimental setup is depicted in \cref{fig:cleave:expsetup}, consisting of a single control loop deployed across a pair of consumer-grade computing devices.
We employ a Raspberry Pi 4B single-board computer together with a general-purpose \verb|x86_64|-based workstation.
These devices are connected through an 802.11n WiFi link.
For ease of deployment and reparametrization, we deploy our CLEAVE control loop as a pair of Docker~\cite{merkel2014docker} containers\todo{add link to image?}; one for the plant, running on the Rasberry Pi device, and one for the Controller, running on the workstation. 
These instances are then connected through an overlay network configured through Docker.

\begin{figure}
    \centering
    \includegraphics[width=.95\columnwidth]{images/CLEAVE_experiment_setup}
    \caption{Experimental setup. Containerized versions of the core CLEAVE emulation components are deployed inside a Docker Overlay Network spanning a Raspberry Pi 4B Plant host connected to an x86 Controller host over a 802.11n WiFi link.}\label{fig:cleave:expsetup}
\end{figure}

We chose a two-dimensional inverted pendulum system for the control system deployed inside CLEAVE for the experiments, 
This system was chosen due to its relative simplicity and prevalence in the field of automatic control as one of the fundamental examples of linear control.
The physical system, represented in \cref{fig:invpend}, was implemented using CLEAVE's API and a 2D physics library~\autocite{chipmunk2d,pymunk}.
For the controller, a proportional-differential strategy was employed, implemented using the framework Controller API and the NumPy numeric computation library~\autocite{harris2020array}.

This setup is then used to run a series of experiments with varying parametrization of the controlled system.
Specifically, we modify:
\begin{itemize}
    \item the sampling rate of the Plant state, and set it to \SIlist[list-units=single,list-final-separator={, or }]{5;10;20;40;60;120}{\hertz};
    \item the responsiveness of the Controller, by adding fixed delays of  \SIlist[list-units=single,list-final-separator={, or }]{0;25;50;100;200}{\milli\second} after the processing of each sample.
\end{itemize}

\begin{figure}
    \centering
    \includegraphics[width=.95\columnwidth]{images/inverted_pendulum.png}
    \caption{
        The two-dimensional inverted pendulum system.
        The cart moves on the X-axis, and the pendulum on top of it swings freely.
        The objective of the system is to balance the pendulum vertically through the application of horizontal forces on the cart.
    }\label{fig:invpend}
\end{figure}
